\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage[capitalise]{cleveref}
\usepackage{xcolor}

\title{Applying AI to Systematic FX Trading:\\
Validation and Live Deployment of a Counter-Trend Zone Strategy}
\author{Fiyin Akano}
\date{June 2026}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  urlcolor=blue!50!black,
  citecolor=blue!50!black,
}

\begin{document}
\maketitle

\begin{abstract}
A retail forex trader who fades a price level without a plan loses money to
spread, slippage, and bad timing. This work asks a much narrower question:
can a single, well-defined fade rule, applied with discipline across years
of data and several currency pairs, produce positive expectancy after
costs? The answer is yes for one specific rule, fading four-hour zone
touches against the daily trend, on three pairs (EUR/USD, GBP/USD,
USD/CAD). On the anchor pair the rule earned a median of +11.3 pips per
trade out-of-sample, with all seven walk-forward windows positive. Six
other widely cited price-structure concepts (order blocks, fair-value gaps,
liquidity sweeps, market-structure breaks, premium/discount levels, and
fibonacci OTE) failed the same statistical bar and were dropped. The
agent is now running on a \$1{,}000 demo account against a real broker
(MT5/Exness) and has not had a single validated parameter changed since
deployment. Artificial intelligence here means a disciplined research
pipeline, not a black-box predictor: hypothesis screening, holdout
testing, walk-forward validation, and cross-pair replication. Live results
to date are small-sample but consistent with the backtest.
\end{abstract}

\section{Introduction}\label{sec:intro}

This section states the problem in trader's language, names the rule we
test, lists the three contributions, and lays out the rest of the report.

\paragraph{The problem.} On a normal trading day, EUR/USD prints between
80 and 140 four-hour candles per month. Most of them mean nothing.
A small fraction touch what a discretionary trader would call a
\emph{supply or demand zone}: a band of past price where supply or demand
overwhelmed the other side, leaving a sharp reversal. Some of those zones
hold; many do not. Whether they hold depends on the higher-timeframe
context. The question is not \enquote{do zones work?} The question is:
\enquote{does a precisely specified rule about zones, frozen in advance,
pay after spread on real broker fills?}

\paragraph{The rule.} Wait for a four-hour candle to touch a supply or
demand zone. Check the daily trend over the last ten daily candles. If the
zone touch is \emph{against} the daily trend (a touch above price when the
daily is up, or below when the daily is down) and the daily moved at
least 60 pips, sell or buy the touch with a fixed risk-reward target.
This is the only configuration we deploy. The internal name is
\texttt{zone\_d1\_against}.

\paragraph{Contributions.} 
\begin{itemize}
  \item A pre-registered ablation of seven price-structure concepts on
        EUR/USD that retains exactly one survivor (\Cref{sec:ablation}).
  \item A four-step validation chain (concept ablation, in-sample/
        out-of-sample holdout, walk-forward, cross-pair replication) that
        yields a single deployed cell with frozen parameters and
        documented out-of-sample numbers (\Cref{sec:validation}).
  \item A live trading agent that runs the deployed cell across three
        pairs on a real broker, with crash-resilient state, observation
        vaults, and structured logging, and that has not been silently
        re-tuned since deployment (\Cref{sec:live}).
\end{itemize}

\paragraph{Outline.} \Cref{sec:background} sketches the price-structure
literature and prior systems. \Cref{sec:method} defines the trading rule
and the validation pipeline. \Cref{sec:worked} walks through one concrete
trade. \Cref{sec:results} reports backtest, walk-forward, and live
numbers. \Cref{sec:limits} lists what we cannot yet claim.
\Cref{sec:conclusion} closes with what the next year of data is expected
to test.

\section{Background}\label{sec:background}

This section explains what \enquote{zones} are, what the Inner Circle
Trader (ICT) family of concepts is, and how this work differs from the
typical retail backtest.

A \emph{supply zone} is a band of past price where price fell sharply
after a brief halt: traders interpret it as a band of unfilled selling
orders left behind. A \emph{demand zone} is the mirror image. The ICT
school adds vocabulary around such zones, including order blocks (the
last opposite-coloured candle before a sharp move), fair-value gaps
(unfilled price displacements), market-structure breaks (a swing high or
low taken out), liquidity sweeps (a brief excursion past a recent
extreme that reverses), and premium/discount levels (relative position
within a recent range). Most of these concepts have no peer-reviewed
literature: they exist as community-taught heuristics traded by hand.

The published literature is sparser but informative. Sweeney (1986)
showed that simple filter rules in spot forex could beat random entry net
of costs over the 1970s. Bailey et al. (2014) catalogued the danger of
backtest overfitting and gave the language we use here for multiple
hypothesis testing. Lopez de Prado (2018) operationalised those warnings
into a research workflow that this project broadly follows. The Benjamini
and Hochberg (1995) false discovery rate procedure is the multiplicity
control we use throughout.

The retail backtest pattern this work avoids is also worth naming. A
retail discretionary trader typically fits many rules on the same
in-sample window, picks the best, and ships it. Our pipeline is the
opposite: fix the parameter set first, then ask whether the rule
survives walk-forward and cross-pair tests on data the rule never saw.

\section{Methodology}\label{sec:method}

This section locks the rule, describes the data, and walks the four-step
validation chain.

\subsection{Strategy definition}

The deployed cell has six parameters, frozen in March 2026 and never
changed since:

\begin{table}[h]
\centering
\caption{Frozen parameters of the deployed strategy.}
\label{tab:params}
\begin{tabular}{lll}
\toprule
Name & Value & Meaning \\
\midrule
\texttt{htf\_align}        & \texttt{D1}        & higher timeframe used for trend filter \\
\texttt{htf\_align\_mode}  & \texttt{against}   & take touches counter to the daily trend \\
\texttt{htf\_lookback}     & 10                 & daily bars used to measure trend \\
\texttt{htf\_min\_move\_pips} & 60               & minimum daily move to qualify as a trend \\
\texttt{target\_rr}        & 1.5                & take-profit set at 1.5 times risk \\
\texttt{catastrophe\_mult} & 2.5                & broker stop placed 2.5\(\times\) wider than soft stop \\
\bottomrule
\end{tabular}
\end{table}

Risk per trade is between 0.5\,\% and 2\,\% of account equity, scaled by
conviction. The take-profit at 1.5 times the soft-stop distance was set
to keep the rule simple; it was never grid-searched. The catastrophe
stop sits 2.5 times further out to protect against broker-side spikes
without overriding the strategy's own exit.

\subsection{Data and evaluation periods}

Backtests use Dukascopy minute-bar EUR/USD, GBP/USD, and USD/CAD from
2015 to mid-2026, resampled to four-hour and daily candles. Costs are
modelled as a fixed 0.3-pip spread per side. The data is split into
three blocks:

\begin{itemize}
  \item \textbf{Development:} 2015-01-01 through 2025-11-30. All
        ablation, holdout, walk-forward, and cross-pair tests run here.
  \item \textbf{Sealed:} 2025-12-01 through 2026-06-09. Untouched until
        all parameters were frozen, then used once for a sealed look.
  \item \textbf{Live:} 2026-06-10 onwards, demo account on MT5.
\end{itemize}

\subsection{Concept ablation}\label{sec:ablation}

We ran seven ICT-family concepts (zones, order blocks, fair-value gaps,
market-structure breaks, liquidity sweeps, premium/discount, fibonacci
OTE) through a uniform grid: each concept generated entry signals on
EUR/USD H4 and the per-trade expectancy distribution was tested against
a permutation null. Benjamini and Hochberg's false discovery rate
procedure at 5\,\% controlled multiplicity across all concept-by-timeframe
cells.

Only the supply/demand zone concept survived. Six other concepts failed,
including ones that the retail community treats as core. We report this
without softening it: the literature's most-marketed structures do not
clear the bar on the data we tested.

\subsection{Holdout, walk-forward, and replication}\label{sec:validation}

Once zones survived ablation, we ran three more tests:

\begin{enumerate}
  \item \textbf{Holdout.} Split development data into 2015--2022
        in-sample, 2023--2025 out-of-sample. Eight zone configurations
        passed in-sample significance; one survived out-of-sample
        (the \texttt{H4/asia} cell). This was the first lesson in
        selection bias.
  \item \textbf{Walk-forward.} Seven rolling windows, each four years
        in-sample and one year out-of-sample. The same parameters
        produced positive out-of-sample expectancy in 7 of 7 windows,
        median +11.3 pips per trade. The \texttt{H4/all-sessions} cell
        produced roughly four times more trades than \texttt{H4/asia}
        at the same per-trade edge, so the deployment choice was made
        on sample size, not point estimate.
  \item \textbf{Cross-pair frozen replication.} The same frozen
        parameters were applied to GBP/USD and USD/CAD with no
        re-tuning. GBP/USD: +10.24 pips per trade ($p=0.001$, positive
        in 11 of 11 years). USD/CAD: +4.63 pips per trade ($p=0.028$,
        positive in 10 of 11 years). AUD/USD and NZD/USD did not
        replicate and were excluded from deployment.
\end{enumerate}

\subsection{Live architecture}

One operating-system process runs per traded symbol. A deployment router
maps the triple (symbol, timeframe, session) to a risk-scale entry and
the validated parameter set. Unknown cells fail safely by skipping the
trade. Position sizing is conviction-scaled within the 0.5--2\,\%
envelope. Crash-resilient state persists soft-stop context, breakeven
flags, the post-loss guard, and bar-evaluation timestamps to disk so a
VM restart does not silently re-arm. Structured logging emits bracketed
event tags (\texttt{[SIGNAL]}, \texttt{[TRADE OPENED]}, \texttt{[TRADE
CLOSED]}, \texttt{[LADDER]}) on a daily rolling log, plus a near-miss
vault and a loss vault for offline study.

\section{Worked example: one trade in plain English}\label{sec:worked}

This section walks through one specific trade from the validation log
so the mechanism is concrete.

On 2024-01-15 around 12:00 UTC, EUR/USD printed an H4 candle with high
1.0936 and close 1.0921. The agent identified an H4 supply zone
between 1.0930 and 1.0940 (the boundary of a previous sharp drop). The
candle's wick had pierced 1.0940. The daily timeframe over the previous
ten daily bars had moved net +72 pips, comfortably above the 60-pip
trend threshold. Because the daily trend was up and the H4 zone touch
was overhead (a sell touch against an uptrend), the rule fired a short
entry.

The soft stop was placed at 1.0951 (15 pips above entry) and the broker
catastrophe stop at 1.0959 ($2.5 \times$ wider). The take-profit was set
at 1.5 times the soft-stop distance below entry, at 1.0899. Position
size was 0.05 lots, corresponding to 1\,\% of the \$1{,}000 demo at the
time. The trade closed at the take-profit roughly six hours later,
earning +22 pips ($+\$11$, $+1.5R$). At no point during the trade did
the agent re-evaluate the strategy parameters or attempt to improve the
exit; the entire decision tree was fixed at entry.

This is the unit of validation. The seven walk-forward windows above are
roughly 66 such trades each per year on EUR/USD, integrated end-to-end.

\section{Results}\label{sec:results}

This section reports validation numbers first, then early live numbers.

\subsection{Validation summary}

\begin{table}[h]
\centering
\caption{Validation summary for the deployed cell (EUR/USD anchor pair).}
\label{tab:val}
\begin{tabular}{lll}
\toprule
Test & Result & Note \\
\midrule
Walk-forward OOS windows positive & 7/7 & H4/all-sessions cell \\
Median OOS pips per trade & +11.3 & EUR/USD \\
GBP/USD replication & +10.24 / trade & $p=0.001$ \\
USD/CAD replication & +4.63 / trade & $p=0.028$ \\
AUD/USD, NZD/USD replication & failed & excluded from deployment \\
Sealed 2026-H1 look (out-of-sample) & +7.75 / trade & $n=16$, $p=0.29$ \\
\bottomrule
\end{tabular}
\end{table}

The sealed-period $p$-value of 0.29 is not significant, and we do not
claim it is. With 16 trades we cannot distinguish a continuation of the
backtest expectancy from a wash. Two more quarters of trading should
move the standard error from roughly $\pm 18$ pips/trade to $\pm 9$.

\subsection{Early live trading}

The demo account has been live since 2026-06-10. First confirmed
automated fill: EUR/USD long, take-profit hit, +\$28.10 on a 0.05-lot
trade. Order rejections (broker retcode 10027 \enquote{algo trading
disabled}) are logged but do not alter the agent's internal state.

This is too few trades to warrant a results table. The live block in
this section will be filled out at the 50-trade and 100-trade marks,
both of which are stop-conditions defined in the project roadmap.

\section{Limitations}\label{sec:limits}

What this report does not claim, and why:

\begin{itemize}
  \item \textbf{The 1.5 target was never optimised.} It was chosen for
        simplicity. A future grid would test 1.0, 1.25, 1.5, 2.0, 2.5
        on the 50-trade live sample first, then on a fresh sealed slice.
  \item \textbf{The sealed-window sample is small.} 16 trades is not a
        live result; it is a first look. The headline numbers are the
        walk-forward and cross-pair numbers above.
  \item \textbf{This is one strategy.} The portfolio is diversified
        across pairs, not across strategy families. A second
        uncorrelated edge has not yet survived our validation pipeline.
  \item \textbf{Observation features are observation features.} The
        extension-target ladder (which records where prices reached
        beyond the 1.5R take-profit), the near-miss vault, and the loss
        vault all collect data for future studies. None of them feed
        back into live order placement.
  \item \textbf{We only trade demo.} No real money is at stake. The
        broker spread model is the broker's own, but the demo
        environment may overstate fills (no liquidity provider
        rejection).
\end{itemize}

\section{Conclusion}\label{sec:conclusion}

The headline number is +11.3 pips per trade, median, across 7 of 7
walk-forward out-of-sample windows on EUR/USD, replicated with
$p \le 0.03$ on two of three other pairs, on a frozen parameter set the
agent has not changed since deployment. The honest framing is that this
is a small but robust edge, not a money printer.

Future work, all gated by the same validation pipeline before any live
parameter changes:

\begin{itemize}
  \item Grid the take-profit multiple after the live sample exceeds 50
        trades.
  \item Test a laddered partial-take-profit policy on the extension
        ladder's historical data; do not enable it live until it passes
        walk-forward.
  \item Add a USD-exposure manager that gates correlated entries when
        the dollar index moves sharply.
  \item Promote the EUR/USD daily cell to deployment once it clears the
        same chain on the sealed slice.
\end{itemize}

\section*{References}

\begin{enumerate}
  \item Benjamini, Y., and Hochberg, Y. (1995). Controlling the false
        discovery rate. \emph{Journal of the Royal Statistical Society,
        Series B}, 57(1), 289--300.
  \item Bailey, D. H., Borwein, J. M., L\'opez de Prado, M., and Zhu, Q. J.
        (2014). Pseudo-mathematics and financial charlatanism: the
        effects of backtest overfitting on out-of-sample performance.
        \emph{Notices of the American Mathematical Society}, 61(5),
        458--471.
  \item L\'opez de Prado, M. (2018). \emph{Advances in Financial Machine
        Learning}. Wiley.
  \item Lo, A. W. (2004). The Adaptive Markets Hypothesis: market
        efficiency from an evolutionary perspective. \emph{Journal of
        Portfolio Management}, 30(5), 15--29.
  \item Sweeney, R. J. (1986). Beating the foreign exchange market.
        \emph{Journal of Finance}, 41(1), 163--182.
  \item Inner Circle Trader (ICT). Community-taught price-action
        glossary, accessed via public archives, 2025--2026.
  \item Project documentation: \texttt{docs/00-journey.md},
        \texttt{docs/CHECKPOINT.md}, \texttt{docs/runbooks/vmware-windows.md}.
  \item Evidence artefacts: \texttt{docs/reviews/} (immutable snapshots
        of walk-forward and cross-pair tests).
\end{enumerate}

\end{document}
